Effectiveness of Instant Feedback on Developer Navigation Practices

### 2013-09-01 paper_id1.tex
interim updates

After classifying the commands we assigned points to the commands in the same configuration file.  Structured navigation commands received 1-10 points each while unstructured navigation commands received 0 points.  Assigning negative points while possible, is discouraged as this may demotivate folks who do not have detailed knowledge of how the study is designed.  When assigning points, make sure you recognize the relationship between points and levels such that "leveling up" happens at a frequent enough pace to make that motivating for users during the study.  Ideally users would level up after a working day or two worth of activity initially extending to a week or even a month following an exponential curve \cite{gamification}.  Designing the points and levels such that above average developers can complete all levels during the study period is ideal.

\subsection{Study Intervention and Participants}
Part of the study can include a communication mechanism that informs the users about the study or even provides training relevant to the objective of the study via a web site.  We discuss this as the "intervention" that expresses the practice that we wish developers to adopt.  For this study we chose to make a nice looking graphic that users click on to see information about navigation commands, install the Sando search tool, and learn about the Blaze game.  The site uses a combination of video and blog-post style information easily accessible to the developer.  The communication site is rated as part of our study completion survey.

Timing the study communication is part of the consideration for a good study design.  We chose to deploy the Blaze tool without communicating the study purpose for a baseline period, then the tool would pop-up a display with the information on the communication site asking the developers to consume the information and begin using the practices.   This allows us to measure changes in the developer's practices over time during the study.  %Other intervention approaches may be used, however, you should take care to understand how the method and timing may affect the measurement of the practices.

We conducted this study with developers working in an R\&D facility in India on a large industrial software system.  Developers were asked to volunteer for the pilot by their management, however, managers did not know who participated in the study.  Keeping managers out of the loop is intended to reduce concerns among the developers that the data may be used for purposes other then the intended study.  By selecting an intact team, we hope to leverage aspects of motivation such as the Leaderboard to spur feelings of competition among teammates.

@@

Murphy-Hill, Parnin, and Black \cite{V:MurphyHill2012How} use the Mylyn Monitor to explore whether or not developers use the automated refactoring tools present in Eclipse.  They look for specific refactoring commands in Eclipse and determine the amount of time developers use tools versus hand refactoring the code.

Robillard, Coelho and Murphy explore hypotheses around how developers can be more effective at performing a maintenance task \cite{wbsnipes:Robillard2004How}.  Key conclusions are that developers are more more successful finding and fixing bugs when they create a detailed plan for implementing a change, use structured navigation with keyword search or cross-reference search, and only review methods once during their search.

>>>>

After classifying the commands we assigned points to the commands in the same configuration file.  Structured navigation commands received 1-10 points each while unstructured navigation commands received 0 points.  Assigning negative points while possible, is discouraged as this may demotivate folks who do not have detailed knowledge of how the study is designed.  When assigning points, make sure you recognize the relationship between points and levels such that "leveling up" happens at a frequent enough pace to make that motivating for users during the study.  Ideally users would level up after a working day or two worth of activity initially extending to a week or even a month following an exponential curve \cite{}.  Designing the points and levels such that above average developers can complete all levels during the study period is ideal.

\subsection{Study Intervention and Participants}

After selecting the practice to focus on, we needed to select an intervention method.  The term intervention, indicates that we are seeking to influence developers to perform a practice previously defined by software engineering community.  Intervention involves communication about the practice in a way that motivates adoption \cite{}.  Interventions can include presentations, one-on-one discussions, distance education mechanisms such as webinars, and word of mouth recommendations.  Part of the research objective is to determine whether we can acheive adoption of a practice with alternative interventions that are more deliberate than word of mouth and perhaps more scalable than webinars or presentations.  Utilizing the automation capabilities of Blaze, we chose to automatically pop-up a window with training information on structured navigation practices and tools. For this study we chose to make a nice looking graphic that users click on to see information about navigation commands, obtain and install the Sando search tool, and learn about the Blaze game.  The site uses a combination of video and blog-post style information easily accessible to the developer.  The usefulness of the communication site is rated as part of our study completion survey.

The study established three time periods, a prelinary period to establish a baseline, a period where information is communicated but no immediate feedback or points are displayed, and a period where points are displayed and all information is available.  We chose to deploy the Blaze tool without communicating the study purpose for the baseline period so that an accurate picture of the baseline could be captured by the tool.  Then the tool would pop-up a display with the information on the communication web site asking the developers to consume the information and begin using the practices.   Finally the display of points fo accumulated by the developer provides the instant feedback mechanism.

We conducted this study with an intact team of developers working in an R\&D facility in India on a large industrial software system.  Developers were asked to volunteer for the pilot by their management, however, managers did not know who participated in the study.  We intended to keep managers out of the loop to reduce concerns among the developers that the data may be used for purposes other then the intended study.  By selecting an intact team, we hope to leverage aspects of motivation such as the Leaderboard to spur feelings of competition among teammates.

@@

Murphy-Hill, Parnin, and Black \cite{V:MurphyHill2012How} use the Mylyn Monitor to explore whether or not developers use the automated refactoring tools present in Eclipse.  They look for specific refactoring commands in Eclipse and determine the amount of time developers use tools versus hand refactoring the code.  For this study we build a tool that captures the use of all commands within the Visual Studio environment including refactoring, navigation, and edit commands.

Robillard, Coelho and Murphy explore hypotheses around how developers can be more effective at performing a maintenance task \cite{wbsnipes:Robillard2004How}.  Key conclusions are that developers are more more successful finding and fixing bugs when they create a detailed plan for implementing a change, use structured navigation with keyword search or cross-reference search, and only review methods once during their search.  This builds on this work by testing methods for increasing the use of structured navigation in developers everyday practice.

Singer and Schneider demonstrate the use of a message board and points for encouraging students to increase their frequency of commits to the source code repository.\cite{Singer2012It}  The communication mechanism enabled students to see each others' progress and resulted in more frequent commits than baseline.  Participants valued the communication and collaboration aspect and some valued the competition enough to change their opinion on optimum commit frequency.  The subsequent thesis by Singer \cite{Singer2013a} describes results of an experiment conducted across two iterations of a class where active feedback on commits was deployed to one course and another course served as the control group where commit frequency was simply monitored.  Results show an increase in the frequency of commits at a statistically significant level.  This study uses finer grained instrumentation to identify practices being used between commits such as structured navigation.

The Pro Metrics (PROM) tool provides a framework for collecting data for further analysis from tools used by developers.\cite{Coman2009Casestudy}  It provides a flexible data model and a plug-in architecture to facilitate collection from different data sources.  Studies conducted using prom include a series of studies by it's creators on trends in time spent Pair Programming \cite{Coman2008Investigating}, benefits of refactoring on productivity \cite{Moser2008Case}, impact of refactoring on re-usability \cite{Moser2006Does}, and prediction of effort \cite{Abrahamsson2007Effort}.  The PROM tool and studies that use it track the time spent editing files and methods from the IDE and code metrics obtained trough source code analysis.  Applying these data to specific research questions such as whether refectoring improves productivity \cite{Moser2008Case} shows the utility of combining automated effort data with code metric data.  Taking a different approach to focus on techniques, our study used more detailed events captured from the IDE  to detect how the user was navigating through the code in addition to knowing how long the editor is open for a particular file.
